---
title: "Lecture_9_markdown"
author: "Cody Ocheltree"
date: "5/2/2018"
output: html_document
keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's load in files needed for the class and test git stuff from the future

```{r}
Wisconsin_cancer <- read.csv("~/Downloads/WisconsinCancer.csv")
new_samples <- read.csv("~/Downloads/new_samples.csv")
```

```{r}
head(Wisconsin_cancer)
table(Wisconsin_cancer$diagnosis)
```

Let's make a data table for the Wisconsin data table including only the numeric values

```{r}
#Wisconsin_df <- Wisconsin_cancer[ , 3:33]
#rownames(Wisconsin_df) <- Wisconsin_cancer$id
#View(Wisconsin_df)
```

Convert the diagnoses from Wisconsin_cancer into a numeric evaluation

```{r}
#diagnosis <- as.numeric(Wisconsin_cancer$diagnosis == "M")
#View(diagnosis)
#table(diagnosis)
```

Let's use PCA. Let's check the mean and standard deviation to determine if the data should be scaled.

```{r}
# Check column means and standard deviations
#plot(colMeans(Wisconsin_df), type = "o")

#plot(apply(Wisconsin_df,2,sd), type = "o")
```

Data needs to be scaled

```{r}
# Wisconsin_pca <- prcomp(Wisconsin_df, scale = T)
# Wisconsin_pca
```

A column in the data frame is causing problems - it lacks any data. Let's remove the column.

```{r}
better_wisconsin_df <- Wisconsin_cancer[ , 3:32]

better_wisconsin_pca <- prcomp(better_wisconsin_df, scale = T)
summary(better_wisconsin_pca)
```

```{r}
plot(better_wisconsin_pca$x[,1], better_wisconsin_pca$x[,2], col=diagnosis+1)
```

```{r}
# Scree plot to determine where the elbow is between PCA's performed (variance)

pca_variance <- better_wisconsin_pca$sdev^2
pve <- pca_variance / sum(pca_variance)

plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
barplot(pve, names.arg=paste("PC", 1:length(pve)), las=2, axes = F, ylab = "Proportion of variance")
axis(2, at=pve, labels=round(pve,2)*100)
```

Section 3 of the mini-project - Hierarchical clustering of a dataset

```{r}
# scaling the dataset for clustering

data_scaled <- scale(better_wisconsin_df)
data_distance <- dist(data_scaled)
wisconsin_hclust <- hclust(data_distance, method = "complete")
plot(wisconsin_hclust)
```

```{r}
wisconsin_hclust_clusters <- cutree(wisconsin_hclust, h=19)
sum(wisconsin_hclust_clusters)
```

Compare the analysis from the hierarchical clustering to that of the diagnosis from the data

```{r}
table(diagnosis, wisconsin_hclust_clusters)
```

This clustering method seems more accurate in binning malignant and benign patients

```{r}
wisconsin_hclust_clusters2 <- cutree(wisconsin_hclust, h=21)
table(diagnosis, wisconsin_hclust_clusters2)
```

Section 5

```{r}
wisconsin_pca_hclust <- hclust(dist(better_wisconsin_pca$x[ , 1:3]), method = "ward.D2")
plot(wisconsin_pca_hclust)

# Trim this clustering into 4 clusters

wisconsin_pca_hclust_clusters <- cutree(wisconsin_pca_hclust, k=2)
table(wisconsin_pca_hclust_clusters)
```


Other stuff - predicting cancer malignancies

```{r}
npc <- predict(better_wisconsin_pca, newdata = new_samples)
plot(better_wisconsin_pca$x[,1:2], col = wisconsin_hclust_clusters2)
points(npc[,1], npc[,2], col=c("purple", "blue"), pch = 16, cex=3)
```

